Current Models

### 2020-01-16 models.Rmd
adding gpedm portalcasting 0.17.0

We currently analyze and forecast rodent data at Portal using five models:

@@

>>>>

We currently analyze and forecast rodent data at Portal using ten models:

@@

## GPEDM

GPEDM (Gaussian processes using Empirical Dynamic Modeling) is a state-space reconstruction model adapted for forecasting and fit to the interpolated data at the composite (full site and just control plots) spatial level and both the composite (community) and the articulated (species) ecological levels. . The method uses time-delay embedding to reconstruct a state-space for the dynamics underlying a time series [@Packard1980, @Takens1981]. The forecast function is approximate using Gaussian processes.

As with `simplexEDM()`, many of the parameters are fit automatically, such as the length-scale, and variance parameters (see `rEDM::block_gp()` for details). One exception is the dimension of the time-delay embedding. Here, the embedding dimension ($E$) is selected as the value (between `1` and the `max_E` argument to `GPEDM()`) that minimizes the mean absolute error over the in-sample portion of the data.
